$i$-UIP Clause Learning

### 2020-02-06 paper.tex
describe first part of the experiment
update marco and package user
added marco's for comments.
make change to optimization 1
add length comparsion and proof size graphs
added author name
first fb changes

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{graphicx}

\begin{document}
%
\title{Contribution Title\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\input{marco}
\maketitle              % typeset the header of the contribution
%
\input{sections/abstract.tex}
\input{sections/introduction.tex}
\input{sections/preliminaries.tex}
\input{sections/i_uip.tex}
\input{sections/evaluation.tex}
%
\begin{thebibliography}{8}

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

>>>>

\documentclass[runningheads]{llncs}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage[ruled,linesnumbered,noend,oldcommands]{algorithm2e}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{theorem}
\usepackage[T1,hyphens]{url}

\begin{document}
\title{$i$-UIP Clause Learning}
\author{Nick Feng \and Fahiem Bacchus}
\authorrunning{N. Feng and F. Bacchus}
\institute{Department of Computer Science, University of Toronto, Canada, 
  \email{\{nfeng,fbacchus\}@cs.toronto.edu}}

\newcommand{\sat}{SAT\xspace}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\inputF}{\ensuremath{F}}
\newcommand{\oneUIPClause}{\ensuremath{C_{1}}}
\newcommand{\iUIPClause}{\ensuremath{C_{i}}}
\newcommand{\assertionTrail}{\ensuremath{\phi}}
\newcommand{\LBD}{\text{LBD}}
\newcommand{\GAP}{\text{Gap}}
\newcommand{\IUIP}{\textbf{$i$-UIP}}
\newcommand{\IUIPPURE}{\textbf{Pure-i-UIP}}
\newcommand{\IUIPMIN}{\textbf{Min-i-UIP}}
\newcommand{\IUIPGreedy}{\textbf{i-UIP-Greedy}}
\newcommand{\IUIPActive}{\textbf{i-UIP-Inclusive}}
\newcommand{\IUIPDist}{\textbf{i-UIP-Exclusive}}
\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
\newcommand{\expSAT}{\textit{expMaple\_CM\_GCBumpOnlyLRB} }
\newcommand{\expSATShort}{\textit{expMaple} }
\newcommand{\MapleIUIPPURE}{\text{Maple-\IUIPPURE}}
\newcommand{\MapleIUIMIN}{\text{Maple-\IUIPMIN}}
\newcommand{\MapleEightShort}{\textit{MapleCB}}
\newcommand\Set[2]{\{\,#1\mid#2\,\}}
\newcommand\Array[2]{[\,#1\mid#2\,]}
\newcommand\target{\textbf{Succ_t}}
\newcommand\trail{\ensuremath{\mathcal{T}}}
\newcommand\trailIdx[1]{\ensuremath{\iota(#1)}}
\newcommand\var{\text{var}}
\newcommand\true{\textsc{true}\xspace}
\newcommand\false{\textsc{false}\xspace}
\newcommand\reason{\textit{reason}}
\setlength{\theorempreskipamount}{3pt}
\setlength{\theorempostskipamount}{3pt}
\newtheorem{Obs}{Observation}
\newtheorem{Cor}{Corollary}
\newtheorem{defn}{Definition}
\newtheorem{thm}{Theorem}

\newcommand{\whitebox}{\raisebox{.5ex}{\fbox{\hspace*{.2ex}}}}

%Comments
\newcommand{\nf}[1]{\textcolor{red}{#1}}
\newcommand{\fb}[1]{\textcolor{blue}{#1}}

\maketitle              % typeset the header of the contribution
% 
\begin{abstract}
\end{abstract}

\section{Introduction}
Clause learning is an essential technique in \sat solvers. And there
is good empirical evidence that clause learning is, in fact, the most
important technique used in modern \sat solvers
\cite{DBLP:conf/sat/KatebiSS11}. In early \sat research a number of
different clause learning techniques were proposed
\cite{DBLP:conf/iccad/ZhangMMM01,DBLP:conf/iccad/SilvaS96,DBLP:journals/tc/Marques-SilvaS99,DBLP:conf/aaai/BayardoS97}.
However, following the revolutionary performance improvements achieved
by the Chaff \sat solver, the field has converged on using the 1-UIP
(first Unique Implication Point) scheme
\cite{DBLP:conf/iccad/ZhangMMM01} employed in Chaff
\cite{DBLP:conf/dac/MoskewiczMZZM01} (as well as other techniques
pioneered in the Chaff solver).\footnote{The idea of UIP clauses was
  first mentioned in \cite{DBLP:journals/tc/Marques-SilvaS99}, and
  1-UIP clauses along with other UIP clauses were learnt and used in
  the earlier GRASP \sat solver.} Since then almost all \sat solvers
have employed the 1-UIP clause learning scheme, along with clause
minimization \cite{DBLP:conf/sat/SorenssonB09}, as their primary
method for learning new clauses.

However, as this early work showed, other clause learning schemes can
be used in SAT solvers without changes to their main data
structures. Furthermore, advances in our understanding allow us to
better understand the potential advantages and disadvantages of these
alternate schemes. In this paper we reexamine some of these previous
clause learning schemes, with a focus on the $i$-UIP schme described
in \cite{DBLP:conf/iccad/ZhangMMM01}. With current ideas we can see
that in its original form $i$-UIP clause learning suffers a
significant disadvantages over 1-UIP clause learning. We present a
number of different methods to impove $i$-UIP clause learning so as to
over come these disadvantages. We show that these different methods
for $i$-UIP clause learning can provide useful performance gains over
the standard 1-UIP scheme.

\section{Background}
We assume the reader has some familiarity with Conjunctive Normal Form
(CNF) formulas, conflict-driven clause learning (CDCL) \sat solvers
\cite{DBLP:series/faia/SilvaLM09}, unit-propagation, and conflict
analysis.

\begin{enumerate}
\item Trail; Reasons and their properties; decision levels, trail order.
\item conflict clause. Trail has the property that conflict exists
    only at the deepest level.
\item Clauses derived by trail resolution. Identified by variables
    appearing on the trail.
\item No point in doing this in non-trail order.
\item a specific case of trivial resolutions.
\item Clause derived by trail resolution iff it is a cut in the
    implication graph.
\item Note that all implemented clause learning techniques use
    sequences of trail resolutions. In particular, they all start with
    a specific conflict and resolve only with reason clauses.
    This includes clause minimization.
\item Asserting clause---one lit at deepest level of trail. Only
    asserting clauses are gauranteed to be new.
\item 1-UIP is asserting clause generated with fewest trail resolution
    steps. Each trail resolution step can only increase the number of
    levels in the clause. So LBD is minimized by 1-UIP clauses.
\item Description of i-uip, all-uip, all decisions (useful in some cases)
\end{enumerate}

CDCL \sat solvers maintain a \textbf{trail} which is a sequence of
literals that have been assigned \true by the solver. Literals newly
assigned \true are placed at the end of the trail, and literals are
always unassigned and removed from the end of the trail first;
i.e. the trail acts as a stack.

Let $\trail$ be the trail. For any literal $\ell$ we let
$\trailIdx{\ell}$ be the index of $\ell$ in $\trail$. That is,
$\trail[\trailIdx{\ell}] = \ell$. If $\ell$ is not in the trail, then
$\trailIdx{\ell} = \varnothing$, and we use the convention that
$\trail[\varnothing] = \varnothing$. The literals on \trail are
sequenced in the order they were assigned \true. Hence
$\trailIdx{\ell} < \trailIdx{\ell'}$ implies that $\ell$ was assigned
\true before $\ell'$. When $\trailIdx{\ell} < \trailIdx{\ell'}$ we say
that $\ell$ \textit{appears above} $\ell'$ on the trail. We use the
convention that $\trailIdx{\ell} < \trailIdx{\ell'}$ whenever $\ell$
is in the trail and $\ell'$ is not; i.e., $i < \varnothing$ for any
index $i$.

The literals on $\trail$ have been assigned \true by the solver by
either (a) having been set as a decisions or (b) having forced by
unit-propagation. If $\ell\in \trail$ was was set by unit-propgation
it will have an associated reason clause $\reason(\ell)$. This clause
has two properties (1) $\ell\in \reason(\ell)$ and (2) for every
$x\in \reason(\ell)$ we have that
$\trailIdx{\lnot x} < \trailIdx{\ell}$. That is, $\reason(\ell)$ is a
clause that implies that $\ell$ must assigned \true because every
other literal in it has been assigned \false above $\ell$ in
$\trail$. Literals $\ell$ set as decsions do not have reason clauses
and we let $\reason(\ell) = \varnothing$.

CDCL \sat solvers perform clause learning when the literals on the
trail falsify a clause $c$. '

\section{Find i-uip Clause}
\label{sec:i-uip}
Classical CDCL SAT solvers analyze each conflict with 1-UIP resolution
learning scheme to derive an asserting clause $\oneUIPClause$. Even
though prior studies shows that further resolutions on $\oneUIPClause$
against the assertion trail $\assertionTrail$ cannot reduce the
clause's literal block distance ($\LBD$), the resolutions could
potentially reduce the clause's size. Clause size is an important
quality measurement because a smaller clause 1) consumes less memory,
2) requires less steps to force a literal and 3) decreases the size of
future conflict clauses.

Inspired by both $\LBD$ and clause size, i-UIP resolution learning
attempts to resolve away literals in $\oneUIPClause$ against the
assertion trail $\assertionTrail$ to minimize the clause size without
increasing the clause's $\LBD$. The goal of i-UIP learning is to find
a clause $\iUIPClause$ whose literals are either the unique
implication points in their respective decision levels or are directly
implied by literals from a foreign decision
level. Alg. ~\ref{alg:i-uip} is a pseudo-code implementation of i-UIP
learning.

\begin{algorithm}[t]
\caption{\IUIP}\label{alg:i-uip}
\begin{algorithmic}[1]
\Require  $\oneUIPClause$ is a valid and minimized 1-UIP clause
\Require  $\assertionTrail$ is a valid assertion trail
\Procedure{\IUIP}{$\oneUIPClause, \assertionTrail$} 
    \State $\iUIPClause \gets \oneUIPClause $ \Comment{initialize i-UIP clause} \label{ln:init}
    \State \code{DecisionLvs} $\gets$ Decision levels in $\oneUIPClause$ in descending order
\For{$i \in \code{DecisionLvs}$} \label{ln:iterDecisionLvls}
    \State $L_{i} \gets \Set{l}{\code{level}(l) = i \wedge l \in \iUIPClause}$
    \While{$|\code{unmarked}(L_i)| > 1$} \label{ln:uip}
        \State $p \gets \text{lit with the highest trail position in } L_{i} $ \label{ln:trailaccess}
        \If{$\exists q \in \code{Reason}(\neg{p}, \assertionTrail) \cdot \code{level}(q) \not\in$ \code{DecisionLvs}} \label{ln:resolvable}
            \If{$\IUIPPURE$}
                \State Unresolve literals at level $i$  \label{ln:unresolve}
                \State Go to the next decision level $i$ \label{ln:abondon}
            \ElsIf{$\IUIPMIN$}
                \State $\code{Mark}(p)$ \label{ln:mark}
            \EndIf
        \Else 
            \State  $\iUIPClause \gets (\oneUIPClause ) \bowtie \code{Reason}(\neg{p}, \assertionTrail)$
           \State  $\code{Update}(L_{i})$
        \EndIf
    \EndWhile
\EndFor \label{ln:endCompleteResolution}
\\
\If{$\IUIPPURE$}\label{ln: 2ndMinStart}   \Comment{$\iUIPClause$ from $\IUIPPURE$ can be minimized}
    \State $\iUIPClause  \gets \textbf{Minimize}(\iUIPClause)$ 
\EndIf\label{ln: 2ndMinEnd}
\\
\If{$|\iUIPClause| < |\oneUIPClause|$} \label{ln:checkSize}
    \State \textbf{return} $\iUIPClause$
\Else
    \State \textbf{return} $\oneUIPClause$
\EndIf  \label{ln:cSizeCheck}
\EndProcedure
\end{algorithmic}
\end{algorithm}

The algorithm $\IUIP$ computes $\iUIPClause$ from (line~\ref{ln:init}
to \ref{ln:endCompleteResolution}), and then returns the smaller
clause between $\oneUIPClause$ and $\iUIPClause$
(line~\ref{ln:checkSize} to \ref{ln:cSizeCheck}).  The algorithm first
initializes $\iUIPClause$ with the minimized\cite{}
$\oneUIPClause$. Next, the algorithm iterate through the decision
levels of $\iUIPClause$ in descending order
(line~\ref{ln:iterDecisionLvls}) and tries to find the unique
implication point in each level (line~\ref{ln:uip}). Since $\IUIP$
needs to preserve $\iUIPClause$'s $\LBD$ during resolutions, before
resolving away a target literal $p$, the algorithm preemptively checks
whether the reasoning clause of $\neg{p}$ contains any literal $q$
from an unseen decision level (line~\ref{ln:resolvable}). If such a
literal $q$ exists, then the algorithm cannot find the unique
implication point (UIP) at the current level $i$ because resolving
away $p$ will introduce $q$ into the clause and increases $\LBD$. One
solution is to abandon level $i$ by reverting back to the state before
any literal at level $i$ is resolved away (line~\ref{ln:unresolve}),
and then move on to the next level (line~\ref{ln:abondon}). We denote
$\IUIP$ learning with this solution as $\IUIPPURE$. The final
$\iUIPClause$ produced by $\IUIPPURE$ will contains exactly one
literal for levels where the LBD-persevering UIP exist. However,
$\IUIPPURE$ does not minimize the number of literals in the decision
levels where the LBD-preserving UIP does not exist, and clause
minimization techniques\cite{} can be used to further reduce the
clause size (line~\ref{ln: 2ndMinStart} to \ref{ln: 2ndMinEnd}).

When $\IUIP$ cannot resolve away a literal $p$ without increasing
$\iUIPClause$'s LBD, instead of skipping the entire decision level
$i$, another solution $\IUIPMIN$ simply marks literal $p$ and keep it
in $\iUIPClause$ (line~\ref{ln:mark}). The learning scheme then
continue the resolutions at level $i$. $\IUIPMIN$ will ignore all the
unsolvable literals and find the "local" unique implication point at
every decision level.  The algorithm terminates when there is exactly
one unmarked literal left for each decision level of $\iUIPClause$,
representing the local unique implication points. The clause
$\iUIPClause$ learnt by $\IUIPMIN$ is minimized\footnote{clause size
  cannot be reduced via minimization techniques\cite{}.}, we defer the
proof of minimization to Appendix~\ref{}.

\begin{example}
    Consider the assertion trail
    $\assertionTrail =
    {..a_2,b_2,c_2,d_2..e_5,f_5,g_5,h_6,i_6,j_6,k_6..}$ where literals
    $a_2 ... k_{6}$ are asserted in order with decision level shown as
    the underscript. Let $L_{ext}$ denote the set of literals on
    $\assertionTrail$ outside of decision level 2,5,6 and 10, and
    $l^{*}_{ext}$ be some literal in $L_{ext}$. The reasoning clauses
    involved in building $\assertionTrail$ are:
\begin{enumerate}
  \item[] $C_k = \{\neg{f_5} \vee \neg{j_6} \vee k_6\}$,
  $C_j = \{\neg{f_5} \vee \neg{i_6} \vee j_6\}$, 
  $C_i = \{\neg{e_5} \vee \neg{h_6} \vee i_6\}$,
  $C_g = \{\neg{a_2} \vee \neg{f_5} \vee g_5\}$
  $C_f = \{\neg{l^{*}_{ext}} \vee \neg{e_5} \vee f_5\}$,
  $C_d = \{\neg{b_2} \vee \neg{c_2} \vee d_2\}$,
  $C_c = \{\neg{a_2} \vee \neg{b_2} \vee c_2 \}$,
  $C_b = \{\neg{l^{*}_{ext}} \vee \neg{a_2} \vee b_2\}$
\end{enumerate}
Suppose 1-UIP learning produces clause
$\oneUIPClause = \{\neg{m_{10}} \vee \neg{k_6} \vee \neg{j_6} \vee
\neg{i_6} \vee \neg{h_6} \vee \neg{g_5} \vee \neg{d_2} \vee
\neg{c_2}\}$ where $\neg{k_{10}}$ is the UIP from the conflicting
level. $\IUIP$ first tries to find the UIP for level 6 by resolving
$\oneUIPClause$ with $C_k$, $C_j$ and, $C_i$ in order and produces
clause
$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{g_5} \vee
\neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \}$ where
$\neg{h_6}$ is the UIP.

$\IUIP$ attempts to find the UIP for level 5 by resolving $C_i$ with
$C_g$ and $C_f$ in order. However, resolving with $C_f$ will introduce
$\neg{l^{*}_{ext}}$ into $C_i$ and increase the clause's
LBD. $\IUIPPURE$ undoes all resolutions at level 5 and skips the
entire level, whereas $\IUIPMIN$ only skips the resolution with $C_f$
and tries to find the UIP by ignoring $\neg{f_5}$.  After level 5,
$\IUIPPURE$ preserves the same clause and $\IUIPMIN$ produces
$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{f_5} \vee
\neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \vee \neg{a_2}\}$ where
$\neg{e_5}$ is the UIP when $\neg{f_5}$ is ignored. The same $\IUIP$
procedure is then performed on level 2 to produce the final
$\iUIPClause$.

$\IUIPPURE$ learned
$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{g_5} \vee
\neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \}$ and reduces
the learnt clause size by 1. $\IUIPMIN$ learned
$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{f_5} \vee
\neg{e_5} \vee \neg{b_2} \vee \neg{a_2}\}$ and reduces the learnt
clause size by 2.
\end{example}

The core algorithm of $\IUIP$ as a clause reduction technique is
simple. However, to make the algorithm a practical learning scheme
that is compatible with modern SAT solvers, a number of issues need to
be addressed. The rest of the section details the key optimizations
and augmentations of $\IUIP$ as a practical clause learning scheme.

\subsection{Control i-UIP Learning}
This simple and greedy $\IUIP$ learning scheme can produce
significantly smaller clause. However, when $\IUIP$ does not reduce
the clause size, the cost of the additional resolution steps will hurt
the solver's performance. Since resolution cannot reduce a clause's
$\LBD$, The maximum size reduction from $\IUIP$ is the difference
between the clause's size and $\LBD$, denote as the clause's gap value
($\GAP(\oneUIPClause) = |\oneUIPClause| - \LBD(\oneUIPClause)$). For
an 1-UIP clause with a small $\GAP$, applying $\IUIP$ is unlikely to
achieve cost effective results. Therefore, we propose a heuristic
based approach to to enable and disable $\IUIP$ learning based on
input clause's $\GAP$.

\begin{algorithm}[t]
\caption{Control-$\IUIP$}\label{alg:enableCondition}
\begin{algorithmic}[1]
\Require  $\oneUIPClause$ is a valid 1-UIP clause
\Require  $ t_{gap} \ge 0$ is a dynamically calculated gap threshold
\Procedure{Control-$\IUIP$}{$\oneUIPClause,  t_{gap}$} 
    \State $Gap \gets |\oneUIPClause| - \textit{LBD}(\oneUIPClause)$
     \If{$Gap > t_{gap}$} \label{ln:compareGap}
        \State $\iUIPClause \gets \IUIP(\oneUIPClause, \phi)$
        \State \textcolor{gray}{\textbf{I-UIP-Greedy}($\iUIPClause$ , $\oneUIPClause$)}
               \Comment{additional clause selection policy, see sec~\ref{sec:greedy}}
        \If {$|\iUIPClause| < |\oneUIPClause|$} 
            \State $\text{Succeed} \gets \text{Succeed}+1$ \label{ln:updateS}
        \EndIf
        \State $\text{Attempted} \gets \text{Attempted}+1$ \label{ln:updateA}
     \EndIf
\EndProcedure
\end{algorithmic}
\end{algorithm}

Alg.~\ref{alg:enableCondition} compares the input $\oneUIPClause$'s
$\GAP$ against a floating target threshold $t_{gap}$
(line~\ref{ln:compareGap}). The threshold $t_{gap}$ represent the
expected minimal $\GAP$ required for $\oneUIPClause$ to achieve a
predetermined success rate (80\%) from performing $\IUIP$ learning.

The gap threshold $t_{gap}$ is initialized to 0, and is updated at
every restart based on $\IUIP$'s success rate from the previous
restart interval. More specifically, the algorithm collects the
statistics of the number of $\IUIP$ learning attempted
(line~\ref{ln:updateA} in alg.~\ref{alg:enableCondition}) and the
number of attempts succeed (line~\ref{ln:updateS}) for each restart
interval, and use them to calculate the success rate. If the success
rate is below 80, the threshold $t_{gap}$ is increased to restrict
$\IUIP$for the next restart interval. On the other hand, the threshold
is decreased to encourage more aggressive $\IUIP$ learning for the
next restart interval.

\[
    t_{gap}=
\begin{cases}
    t_{gap} + 1& \text{if } \frac{\text{Succeed}}{\text{Attempted}} < 0.8\\
    max(t_{gap} - 1, 0),              & \text{otherwise}
\end{cases}
\]

\subsection{Early stop i-UIP Learning}
At any point of $\IUIPMIN$ learning, if the number of marked literals
in $\iUIPClause$ (literals which are forced into the clause to
preserve $\LBD$) exceeds the input clause $\oneUIPClause$'s $\GAP$, we
can abort $\IUIP$ learning for $\oneUIPClause$ because the size of the
final $\iUIPClause$ is at least the size of $\oneUIPClause$.  The
early stopping rule prevents solver from wasting time on traversing a
large implication graph when $\IUIP$ has already failed.

The early stopping rule may prematurely stops $\IUIPPURE$ learning
which would otherwise succeed because $\iUIPClause$ can be further
minimized. However, the experiments indicates such false negative
cases are rare, and the early stopping rule generally helps the
learning scheme's performance.

\subsection{Greedy Active Clause Selection} \label{sec:greedy} Even
though $\IUIP$ learning can reduce the size of the learnt clause
$\iUIPClause$, but it may introduce literals with low variable
activity into $\iUIPClause$.  Inactive literals prevents the clause
from being asserted to force literal implication. Therefore, a
practical clause learning scheme should consider both size and
variable activity. We propose an optional extension $\IUIPGreedy$ to
filter out inactive $\oneUIPClause$.

\begin{algorithm}[t]
\caption{$\IUIPGreedy$}\label{alg:greedySelection}
\begin{algorithmic}[1]
\Require  $\iUIPClause$ is a valid i-UIP clause
\Require  $\oneUIPClause$ is a valid 1-UIP clause
\Procedure{$\IUIPGreedy$}{$\iUIPClause, \oneUIPClause$} 
   \If{$|\iUIPClause| < |\oneUIPClause| \wedge (\code{AvgVarAct}(\iUIPClause) > \code{AvgVarAct}(\oneUIPClause)) $} \label{ln:checkSizeAndActivity}
    \State \textbf{return} $\iUIPClause$
\Else
    \State \textbf{return} $\oneUIPClause$
\EndIf  \label{ln:cSizeCheck}
\EndProcedure
\end{algorithmic}
\end{algorithm}

After computing the $\iUIPClause$, $\IUIPGreedy$ compares both the
size and the average variable activity for $\oneUIPClause$ and
$\iUIPClause$ (alg.~\ref{alg:greedySelection} at
line~\ref{ln:checkSizeAndActivity}). The algorithm learns
$\iUIPClause$ if the clause has smaller size and higher average
variable activity.

\subsection{Adjust Variable Activity} \label{sec: varajust} Two
popular branching heuristics for modern SAT solvers are VSIDS and
LBR. Both heuristics increase the variable activity for all literals
involved in resolutions during 1-UIP learning. Since $\IUIP$ extends
1-UIP with deeper resolutions against the trail, the variables
activities for the fresh literals involved in the additional
resolution steps need to be adjusted as well. We purpose two optional
schemes for adjusting variable activities, $\IUIPActive$ and
$\IUIPDist$.

After learning $\oneUIPClause$ from 1-UIP scheme, $\IUIPActive$
collects all literals appear in the the $\IUIP$ clause $\iUIPClause$,
and increase their variable activity uniformly according to the
current branching heuristic if the literals' variable activity have
not yet been bumped during 1-UIP. The scheme does not bump variable
activities for transient literals that are resolved away at
non-conflicting level because, unlike transient literals at
conflicting level, these literals cannot be re-asserted after the
immediate backtracking. Notice that other post-analyze extensions such
as Reason Side Rate (RSR) and Locality\cite{} can be applied after
$\IUIPActive$.

\[ \forall l \in  \iUIPClause \cdot NotBumped(var(l)) \implies  bumpActivity(var(l)) \]

$\IUIPDist$ collects literals appear exclusively in $\iUIPClause$ and
bump their variable activity uniformly. It also find all the literals
in $\oneUIPClause$ that are resolved away during $\IUIP$, and unbump
their variable activity if they have been bumped during 1-UIP. The
unbumped literals are no longer in the learned clause $\iUIPClause$,
keep their variable activity bumped does not help solver to use
$\iUIPClause$.

\[ \forall l_{i} \in  \iUIPClause \setminus \oneUIPClause \cdot bumpActivity(var(l_{i})) 
\]
\[ \forall l_{1} \in  \oneUIPClause \setminus \iUIPClause \cdot unbumpActivity(var(l_{1})) 
\]

\subsection{Integrate with Chronological Backtracking}
In SAT solver with chronological backtracking, literals on the
assertion trail $\assertionTrail$ are not always sorted by decision
levels. This change imposes a challenge to $\IUIP$ learning since the
previous implementation relies on solver's ability to efficiently
access all literals from any decision level in descending trail order
(alg.~\ref{alg:i-uip} line~\ref{ln:trailaccess}).

To mitigate this challenge, we modify the solver to track the precise
trail position of all asserted literals with a single vector. We then
build a priority queue $lit\_Order$ to manage literal's resolution
order. The queue $lit\_Order$ prioritizes literals with higher
decision level, and it favors literal with higher trail position when
decision levels are tied. The order of $lit\_Order$ represents the
correct resolution order of $\IUIP$ because 1) an asserted literal's
decision level is the maximum level among all of its reasoning
literals, and 2) an asserted literal always appears higher in the
trail than its reasoning literal.

\begin{algorithm}[t]
\caption{\IUIP-CB}\label{alg:i-uip-CB}
\begin{algorithmic}[1]
\Require  $\oneUIPClause$ is a valid 1-UIP clause
\Require  $\assertionTrail$ is a valid assertion trail
\Require  $lit\_Order$ is a priority queue
\Procedure{\IUIP-CB}{$\oneUIPClause, \assertionTrail, lit\_Order $} 
    \State ...
    \State $\iUIPClause \gets \oneUIPClause $ \Comment{initialize i-UIP clause}
    \State forall $l \in \oneUIPClause \cdot \code{Enqueu}(lit\_Order, l)$  \label{ln:enqueue}
    \State \code{DecisionLvs} $\gets \Set{\code{level}(l)}{l \in \oneUIPClause}$
    \While{$lit\_Order \neq \emptyset$} \label{ln:pop}
        \State $p \gets \code{dequeu}(lit\_Order, l) $ \label{ln:dequeue}
        \If{$\exists q \in \code{Reason}(\neg{p}, \assertionTrail) \cdot \code{level}(q) \not\in$ \code{DecisionLvs} \\ $\vee p$ is the last reminaing lit in its decision level } \label{ln:resolvable}
          \State Pass
        \Else 
            \State  $\iUIPClause \gets (\oneUIPClause ) \bowtie \code{Reason}(q, \assertionTrail)$
            \State   forall $l \in \code{Reason}(q, \assertionTrail) \cdot \code{Enqueu}(lit\_Order, l)$ \label{ln:reequeue}
        \EndIf
    \EndWhile
\State ...
\EndProcedure
\end{algorithmic}
\end{algorithm}

Alg.~\ref{alg:i-uip-CB} is the pesudo-code implementation of the
augmented $\IUIP$ for chronological backtracking with the priority
queue $lit\_Order$. The algorithm first populates $lit\_Order$ with
all literals in $\oneUIPClause$ (line~\ref{ln:enqueue}), and then
continuously pops literals until the queue is empty
(line~\ref{ln:pop}). When a literal is resolved away, all of its
reasoning literals are added into $lit\_Order$ if they are not already
in the queue (line~\ref{ln:reequeue}). The algorithm will always
terminate because a literal cannot entered the queue twice (guaranteed
by the properties of the trail order) and there are finite amount of
literals on the trail.

\section{Implementation and Experiments}

\subsection{Clause Reduction with $\IUIP$}
To evaluate $\IUIP$'s effectiveness as a clause reduction technique,
we implement $\IUIPPURE$ and $\IUIPMIN$ with Control-I-UIP and Early
Stopping rules in Sec~\ref{sec:i-uip} on top of \text{\MapleBase}
\cite{}, the winner of SAT Race 2015 application track.  We than
compare the performance of the baseline $\MapleBase$ with
$\MapleIUIPPURE$ and $\MapleIUIMIN$ on the full set of benchmarks from
SAT RACE 2019 main track.

The benchmark contains 400 instances divided into two groups of 200,
new and old, representing historical instances and fresh instances in
the 2019 race, respectively. We randomly partition the old group
instances into six partitions of size 30 and one partition of size
20. Each partition is then assigned to a XeonE5-2 CPU node with 16
cores (2 sockets 8 cores and 1 thread) and 96649 MB memory. The new
group is partitioned based on their contributor (e.g. Heule
contributed 22 matrix multiplication instances), and each partition is
assigned to a aforementioned CPU node. To speed up the experiment, we
allow a CPU node to solver at most seven instances concurrently.

Beside solved instances count and PAR-2 score, we additionally measure
the average clause length and clause reduction ratio (both cumulative
and non-cumulative)\footnote{The cumulative reduction ratio is
  obtained through learning all clauses with the target learning
  scheme; Therefore, the reduction is cumulative. The non-cumulative
  reduction ratio is obtained by running the target scheme for
  measurement only (the minimized 1-UIP clause is learned); Therefore,
  the reduction is not cumulative. }  for each instances. For
$\MapleIUIPPURE$ and $\MapleIUIMIN$, we also captures the $\IUIP$
learning attempted rate and success rate.

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
\hline
Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
\hline
$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
\hline
$\MapleIUIPPURE$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
\hline
$\MapleIUIMIN$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPURE$  and $\MapleIUIMIN$ on SAT2019 race main track.
CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, and the values in the brackets are the non-cumulative reduction ratio.}
\label{fig:t1}
\end{figure}

Fig.~\ref{fig:t1} shows that $\IUIPPURE$ ($\IUIPPURE$) solved seven
(five) more instances than the baseline solver with lower PAR-2
scores. $\IUIPPURE$ has marginally lower PAR-2 score than
$\IUIPPURE$. Both $\IUIPPURE$ and $\IUIPMIN$ produce clause with
significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
respectively. Fig.~\ref{fig:len_pdf} shows the probability density
distribution (PDF) of the relative clause size of both i-uip learning
schemes ($\IUIPPURE$ in red and $\IUIPMIN$ in green) for each
instance. $\IUIPPURE$ ($\IUIPMIN$) produces shorter clauses for 77.7\%
(88.5\%) of instances, and average relative reduction from 1-UIP is
16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
clause size of $\IUIPMIN$, $\IUIPPURE$ and 1-UIP, and it shows that
both i-UIP learning schemes in general produce the smaller clauses,
and the size reduction is more significant for instances with large
average 1-UIP clause size. Moreover, $\IUIPMIN$ clauses (indicated in
green) is consistently smaller than $\IUIPPURE$ clauses.
